---
title: "Simulations HW4"
author: "Adil Hayat"
date: "14 April 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(OptionPricing)
set.seed(0)
```

## Generic function to generate paths of Brownian Motion
```{r}
generateGeomBMPaths <- function(S0=S0, r=r, sigma=sigma,Time=Time, N=N, div=0) {
  stockPrice <- numeric()
  deltaTime <- Time/N
  stockPrice[1] <- S0*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rnorm(1)*sqrt(deltaTime))   
  for(k in 2:N){
   stockPrice[k] = stockPrice[k-1]*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rnorm(1)*
                                         sqrt(deltaTime))   
  }
  return(stockPrice)
}
```

## Question 1
```{r}
# Parameters
S0 <- 100
K <- 100
r <- 0.05
V0 <- 0.04
Time <- 1
sampl_size <- 10000
h <- 0.0001
N <- 50
comp_df <- data.frame(matrix(nrow=2,ncol=6))
colnames(comp_df) <- c("alpha","psi","SMCPrice","SMCSE","CMCPrice","CMCSE")
comp_df$alpha <- 0.1
comp_df$psi <- c(0.1,1.0)
```

## Question 1(a)
```{r, cache=TRUE}
deltaTime <- Time/N
for(j in 1:nrow(comp_df)){
  alpha <- comp_df$alpha[j]
  psi <- comp_df$psi[j]
  call_prices <- numeric()
  for(i in 1:sampl_size){
    stockPrice <- numeric()
    variance <- numeric()
    stockPrice[1] <- S0 + r*S0*deltaTime + sqrt(V0)*rnorm(1)*S0*sqrt(deltaTime)
    variance[1] <- V0*exp((alpha-0.5*psi^2)*deltaTime+psi*rnorm(1)*sqrt(deltaTime))
    for(k in 2:N){
      stockPrice[k] <- stockPrice[k-1] + r*S0*deltaTime + 
                       sqrt(variance[k-1])*rnorm(1)*stockPrice[k-1]*sqrt(deltaTime)
      variance[k] <- variance[k-1]*exp((alpha-0.5*psi^2)*deltaTime + 
                       psi*rnorm(1)*sqrt(deltaTime))
    }
    call_prices[i] <- exp(-r*Time)*max(stockPrice[N]-K,0)
  }
  call_price <- mean(call_prices)
  se_call <- sqrt(sum((call_prices - call_price)^2)/(sampl_size*(sampl_size-1)))
  comp_df$SMCPrice[j] <- call_price
  comp_df$SMCSE[j] <- se_call  
}
```

## Question 1(b)
```{r}
for(j in 1:nrow(comp_df)){
  alpha <- comp_df$alpha[j]
  psi <- comp_df$psi[j]
  call_prices <- numeric()
  for(i in 1:sampl_size){
    variances <- generateGeomBMPaths(S0 = V0, r = alpha, sigma = psi, Time = Time, N = N)
    sig <- sqrt(mean(variances))
    call_prices[i] <- as.double(BS_EC(S0 = S0,T = Time, K = K, r = r, sigma = sig)[1])
  }
  call_price <- mean(call_prices)
  se_call <- sqrt(sum((call_prices - call_price)^2)/(sampl_size*(sampl_size-1)))
  comp_df$CMCPrice[j] <- call_price
  comp_df$CMCSE[j] <- se_call  
}
print(comp_df)
```

We observe that the standard errors for the Conditional Monte Carlo much lower than in the case with Standard Monte Carlo simulation.

## Question 2
```{r}
# Parameters
sampl_size <- 1000
alpha <- 0.2
sigma <- 0.1
b <- 0.05
R0 <- 0.04
N <- 50
Time <- 1

generateCIRPath <- function(alpha=alpha, sigma=sigma, b=b, R0=R0, N=N, Time=Time) {
  rate <- numeric()
  deltaTime <- Time/N
  d <- (4*b*alpha)/sigma^2
  lam <- (4*alpha*exp(-alpha*deltaTime))*R0/(sigma^2*(1-exp(-alpha*deltaTime)))
  rate[1] <- sigma^2*(1-exp(-alpha*deltaTime))*rchisq(n = 1, df = d, ncp = lam)/
            (4*alpha)  
  for(k in 2:N){
    lam <- (4*alpha*exp(-alpha*deltaTime))*rate[k-1]/
            (sigma^2*(1-exp(-alpha*deltaTime)))
    rate[k] <- sigma^2*(1-exp(-alpha*deltaTime))*rchisq(n = 1, df = d, ncp = lam)/
               (4*alpha)  
  }
  return(rate)
}
```
## Question 2(a)
```{r}
bondPrices <- numeric()
deltaTime <- Time/N
for(i in 1:sampl_size){
  rate <- generateCIRPath(alpha, sigma, b, R0, N, Time)
  bondPrices[i] <- exp(-deltaTime*sum(rate))
}

bond_price <- mean(bondPrices)
se_bond_price <- sqrt(sum((bondPrices - bond_price)^2)/
                        (sampl_size*(sampl_size-1)))
cat("Bond Price: ", bond_price, "\n")
cat("se(Bond Price): ", se_bond_price, "\n")
```
## Question 2(b)
```{r}
strikeRate <- 0.05
del <- 1/12
L <- 1
capletPrices <- numeric()
for(i in 1:sampl_size){
 rate <- generateCIRPath(alpha, sigma, b, R0, N, Time)
 capletPrices[i] <- L*del*max(rate[N]-strikeRate,0)*exp(-deltaTime*sum(rate))
}

capletPrice <- mean(capletPrices)
se_caplet_price <- sqrt(sum((capletPrices - capletPrice)^2)/
                          (sampl_size*(sampl_size-1)))
cat("Caplet Price: ", capletPrice, "\n")
cat("se(Caplet Price): ", se_caplet_price, "\n")
```

## Question 3
```{r}
# Parameters
S0 <- 90
K <- 100
r <- 0.1
sigma <- 0.25
Time <- 0.2
sampl_size <- 10000
h <- 0.0001
N <- 50
del <- 0.03


# generates paths at S0+h and S0
generateDelPaths <- function(S0=S0, K=K, r=r, sigma=sigma,Time=Time, N=N, h=h, div=0) {
  Sminus <- numeric()
  Splus <- numeric()
  deltaTime <- Time/N
  rand <- rnorm(1)
  Sminus[1] <- (S0)*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rand*sqrt(deltaTime))   
  Splus[1] <- (S0+h)*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rand*sqrt(deltaTime))   
  for(k in 2:N){
    rand <- rnorm(1)
    Sminus[k] <- Sminus[k-1]*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rand*
                                   sqrt(deltaTime))   
    Splus[k] <- Splus[k-1]*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rand*
                                 sqrt(deltaTime))   
  }
  return(list(Splus,Sminus))
}

# generates paths at sig0+h and sig0
generateSigDelPaths <- function(S0=S0, K=K, r=r, sigma=sigma,Time=Time, N=N, h=h, div=0) {
  Sminus <- numeric()
  Splus <- numeric()
  deltaTime <- Time/N
  rand <- rnorm(1)
  Sminus[1] <- (S0)*exp((r-div-0.5*(sigma)^2)*deltaTime+(sigma)*rand*sqrt(deltaTime))   
  Splus[1] <- (S0)*exp((r-div-0.5*(sigma+h)^2)*deltaTime+(sigma+h)*rand*sqrt(deltaTime))   
  for(k in 2:N){
    rand <- rnorm(1)
    Sminus[k] <- Sminus[k-1]*exp((r-div-0.5*(sigma)^2)*deltaTime+(sigma)*
                                   rand*sqrt(deltaTime))   
    Splus[k] <- Splus[k-1]*exp((r-div-0.5*(sigma+h)^2)*deltaTime+(sigma+h)*
                                 rand*sqrt(deltaTime))   
  }
  return(list(Splus,Sminus))
}

```
## Calculation of values
```{r}
res_df <- data.frame(matrix(nrow = 6,ncol=5))
colnames(res_df) <- c("Method", "DeltaEst", "DeltaStdErr", "VegaEst", "VegaStdErr")
res_df$Method <- c("Resimulation estimate","Resimulation with control",
                      "Pathwise estimate", "Pathwise with control",
                      "Likelyhood estimate", "Likelyhood with control")

## Resampling
# calculate deltas
deltas <- numeric()
sexp <- numeric()
for(i in 1:sampl_size){
  pricePaths <- generateDelPaths(S0=S0, K=K, r=r, sigma=sigma,Time=Time, N=N, h=h, div = del)
  callPricePlus <- exp(-r*Time)*ifelse(pricePaths[[1]][N]-K > 0, pricePaths[[1]][N]-K, 0)
  callPriceMinus <- exp(-r*Time)*ifelse(pricePaths[[2]][N]-K > 0, pricePaths[[2]][N]-K, 0)
  deltas[i] <- (callPricePlus - callPriceMinus)/(h)
  sexp[i] <- pricePaths[[2]][N]
}

delta <- mean(deltas)
seDelta <- sqrt(sum((deltas - delta)^2)/(sampl_size*(sampl_size-1)))
res_df$DeltaEst[1] <- delta
res_df$DeltaStdErr[1] <- seDelta

# with control
aEst <- -1*cov(sexp, deltas)/var(sexp)
deltasCont <- deltas + aEst*(sexp-S0*exp((r-del)*Time))
res_df$DeltaEst[2] <- mean(deltasCont)
res_df$DeltaStdErr[2] <- sqrt(sum((deltasCont - mean(deltasCont))^2)/
                                (sampl_size*(sampl_size-1)))

# calculate vegas  
vegas <- numeric()
sexp <- numeric()
for(i in 1:sampl_size){
  pricePaths <- generateSigDelPaths(S0=S0, K=K, r=r, sigma=sigma, Time=Time, 
                                    N=N, h=h, div = del)
  callPricePlus <- exp(-r*Time)*max(pricePaths[[1]][N]-K, 0)
  callPriceMinus <- exp(-r*Time)*max(pricePaths[[2]][N]-K, 0)
  vegas[i] <- (callPricePlus - callPriceMinus)/(h)
  sexp[i] <- pricePaths[[2]][N]
}
vega <- mean(vegas)
seVega <- sqrt(sum((vegas - vega)^2)/(sampl_size*(sampl_size-1)))
res_df$VegaEst[1] <- vega
res_df$VegaStdErr[1] <- seVega
# with control
aEst <- -1*cov(sexp, vegas)/var(sexp)
vegasCont <- vegas + aEst*(sexp-S0*exp((r-del)*Time))
res_df$VegaEst[2] <- mean(vegasCont)
res_df$VegaStdErr[2] <- sqrt(sum((vegasCont - mean(vegasCont))^2)/
                               (sampl_size*(sampl_size-1)))


## Pathwise Estimate
deltas <- numeric()
vegas <- numeric()
for(i in 1:sampl_size){
  pricePaths <- generateGeomBMPaths(S0=S0,r = r, sigma=sigma, Time=Time, N=N, div=del)
  deltas[i] <- exp(-r*Time)*ifelse(pricePaths[N]-K > 0, 1, 0)*pricePaths[N]/S0
  vegas[i] <- exp(-r*Time)*ifelse(pricePaths[N]-K > 0, 1, 0)*
              (log(pricePaths[N]/S0)-(r-del+0.5*sigma^2)*Time)*pricePaths[N]/sigma
  sexp[i] <- pricePaths[N]
}

delta <- mean(deltas)
seDelta <- sqrt(sum((deltas - delta)^2)/(sampl_size*(sampl_size-1)))
res_df$DeltaEst[3] <- delta
res_df$DeltaStdErr[3] <- seDelta
# with final stock price as control
aEst <- -1*cov(sexp, deltas)/var(sexp)
deltasCont <- deltas + aEst*(sexp-S0*exp((r-del)*Time))
res_df$DeltaEst[4] <- mean(deltasCont)
res_df$DeltaStdErr[4] <- sqrt(sum((deltasCont - mean(deltasCont))^2)/
                                (sampl_size*(sampl_size-1)))
# vega
vega <- mean(vegas)
seVega <- sqrt(sum((vegas - vega)^2)/(sampl_size*(sampl_size-1)))
res_df$VegaEst[3] <- vega
res_df$VegaStdErr[3] <- seVega
# with final stock price as control
aEst <- -1*cov(sexp, vegas)/var(sexp)
vegasCont <- vegas + aEst*(sexp-S0*exp((r-del)*Time))
res_df$VegaEst[4] <- mean(vegasCont)
res_df$VegaStdErr[4] <- sqrt(sum((vegasCont - mean(vegasCont))^2)/
                               (sampl_size*(sampl_size-1)))


## Likelyhood Estimate
deltas <- numeric()
vegas <- numeric()
sexp <- numeric()
for(i in 1:sampl_size){
  pricePaths <- generateGeomBMPaths(S0=S0, r=r, sigma=sigma,Time=Time, N=N, div=del)
  deltas[i] <- exp(-r*Time)*max(pricePaths[N]-K,0)*(log(pricePaths[N]/S0)-
              (r-del-0.5*sigma^2)*Time)/(S0*sigma^2*Time)
  d <- (log(pricePaths[N]/S0)-(r-del-0.5*sigma^2)*Time)/(sigma*sqrt(Time))
  deriv <- (log(S0/pricePaths[N]) + (r-del+0.5*sigma^2)*Time)/(sigma^2*sqrt(Time))
  vegas[i] <- exp(-r*Time)*max(pricePaths[N]-K, 0)*(-d*deriv-(1/sigma))
  sexp[i] <- pricePaths[N]
}
# delta
delta <- mean(deltas)
seDelta <- sqrt(sum((deltas - delta)^2)/(sampl_size*(sampl_size-1)))
res_df$DeltaEst[5] <- delta
res_df$DeltaStdErr[5] <- seDelta
# with final stock price as control
aEst <- -1*cov(sexp, deltas)/var(sexp)
deltasCont <- deltas + aEst*(sexp-S0*exp((r-del)*Time))
res_df$DeltaEst[6] <- mean(deltasCont)
res_df$DeltaStdErr[6] <- sqrt(sum((deltasCont - mean(deltasCont))^2)/
                                (sampl_size*(sampl_size-1)))
# vega
vega <- mean(vegas)
seVega <- sqrt(sum((vegas - vega)^2)/(sampl_size*(sampl_size-1)))
res_df$VegaEst[5] <- vega
res_df$VegaStdErr[5] <- seVega
# with final stock price as control
aEst <- -1*cov(sexp, vegas)/var(sexp)
vegasCont <- vegas + aEst*(sexp-S0*exp((r-del)*Time))
res_df$VegaEst[6] <- mean(vegasCont)
res_df$VegaStdErr[6] <- sqrt(sum((vegasCont - mean(vegasCont))^2)/
                               (sampl_size*(sampl_size-1)))
print(res_df)
```


## Question 4
```{r}
# Parameters
S0 <- 95
K <- 100
r <- 0.05
sigma <- 0.2
Time <- 1
sampl_size <- 10000
h <- 0.0001
N <- 50
```

## Question 4(a)
$$P(S_0,K,r,T) = e^{-rT}N(d_{-})$$
          
$$\frac{\partial P(S_0,K,r,T)}{\partial S_0}=e^{-rT}N^{'}(d_{-})\frac{1}{S_0{\sigma}\sqrt{T}}$$
$$Delta(\frac{\partial P}{\partial S_0})= e^{-rT}N^{'}(d_{-})\frac{1}{S_0{\sigma}\sqrt{T}}$$

```{r}
## theoretical value
dminus <- (log(S0/K)+(r-0.5*sigma^2))/(sigma*sqrt(Time))
th_val <- exp(-r*Time)*(dnorm(dminus))/(S0*sigma*Time)
cat("Theoretical Value:", th_val,"\n")
```

## Question 4(b)
## Resimulation method
```{r}
deltas <- numeric()
for(i in 1:sampl_size){
  pricePaths <- generateDelPaths(S0=S0, K=K, r=r, sigma=sigma,Time=Time, N=N, h=h)
  callPricePlus <- exp(-r*Time)*ifelse(pricePaths[[1]][N]-K > 0, 1, 0)
  callPriceMinus <- exp(-r*Time)*ifelse(pricePaths[[2]][N]-K > 0, 1, 0)
  deltas[i] <- (callPricePlus - callPriceMinus)/(h)
}

delta <- mean(deltas)
seDelta <- sqrt(sum((deltas - delta)^2)/(sampl_size*(sampl_size-1)))
cat("Delta: ", delta, "\n")
cat("se(Delta): ", seDelta, "\n")
```
## Question 4(c)
## Likelyhood Method
```{r}
deltas <- numeric()
for(i in 1:sampl_size){
  price <- generateGeomBMPaths(S0=S0, r=r, sigma=sigma,Time=Time, N=N)
  deltas[i] <- exp(-r*Time)*ifelse(price[N]-K > 0, 1, 0)*(log(price[N]/S0) -
               (r-0.5*sigma^2))/(S0*sigma^2*Time)
}

delta <- mean(deltas)
seDelta <- sqrt(sum((deltas - delta)^2)/(sampl_size*(sampl_size-1)))
cat("Delta: ", delta, "\n")
cat("se(Delta): ", seDelta, "\n")
```
