---
title: "Simulations HW5"
author: "Adil Hayat"
date: "19 April 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
```

## Generic function to generate paths of Brownian Motion
```{r}
generateGeomBMPaths <- function(S0=S0, r=r, sigma=sigma,Time=Time, N=N, div=0) {
  stockPrice <- numeric()
  deltaTime <- Time/N
  stockPrice[1] <- S0*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rnorm(1)*sqrt(deltaTime))   
  for(k in 2:N){
   stockPrice[k] = stockPrice[k-1]*exp((r-div-0.5*sigma^2)*deltaTime+sigma*rnorm(1)*
                                         sqrt(deltaTime))   
  }
  return(stockPrice)
}

generateTerminalGeomBMPaths <- function(S0=S0, r=r, sigma=sigma,Time=Time, div=0) {
  stockPrice <- S0*exp((r-div-0.5*sigma^2)*Time+sigma*rnorm(1)*sqrt(Time))   
  return(stockPrice)
}

generateTermGeomBMPathsUnif <- function(S0=S0, r=r, sigma=sigma,Time=Time,
                                        min=0, max=1,div=0) {
  stockPrice <- S0*exp((r-div-0.5*sigma^2)*Time+sigma*
                qnorm(runif(1, min = min, max = max))*sqrt(Time))   
  return(stockPrice)
}

calculateMeanSE <- function(var){
  return(sqrt(sum((var - mean(var))^2)/(length(var)*(length(var)-1))))
}
```

## Question 1
## parameters
```{r paramsq1}
S0 <- c(100,100)
sigma <- c(0.2, 0.2)
Time <- 1
r <- 0.05
K <- 100
sampl_size <- 1000
comp_df <- data.frame(matrix(nrow=3,ncol=3))
colnames(comp_df) <- c("Method","Call_Price","SE_Call_Price")
comp_df$Method <- c("Standard Monte Carlo", "Bivariate Stratification", 
                    "Conditioning on Projection")
```
## 1(a)
```{r, dependson="paramsq1",cache=TRUE}
S1 <- numeric()
S2 <- numeric()
call_prices <- numeric()
for(i in 1:sampl_size){
  S1[i] <- generateTermGeomBMPathsUnif(S0 = S0[1], r = r, sigma = sigma[1],
          Time = Time)
  S2[i] <- generateTermGeomBMPathsUnif(S0 = S0[2], r = r, sigma = sigma[2],
          Time = Time)
  call_prices[i] <- exp(-r*Time)*max( 0.5*(S1[i] + S2[i]) - K, 0)
}

call_price <- mean(call_prices)
se_call <- sqrt(sum((call_prices - call_price)^2)/(sampl_size*(sampl_size-1)))
comp_df$Call_Price[1] <- call_price
comp_df$SE_Call_Price[1] <- se_call
cat("Call Price and Standard Error\n")
cat("Call Price: ", call_price, "\n")
cat("Standard Error: ", se_call)
```
## 1(b)
```{r 1(b), dependson="paramsq1", cache=TRUE}
bin_size <- 10
num_bins <- sampl_size/bin_size
x_bins <- 10
y_bins <- 10  
bin_price <- matrix(nrow = 10,ncol = 10)
bin_var <- matrix(nrow = 10,ncol = 10)

for(k in 1:x_bins){
  for(j in 1:x_bins){
    S1 <- numeric()
    S2 <- numeric()
    call_prices <- numeric()
    for(i in 1:bin_size){
      x_min <- 0.1*(k-1) 
      x_max <- 0.1*(k)
      y_min <- 0.1*(j-1) 
      y_max <- 0.1*(j)
      S1[i] <- generateTermGeomBMPathsUnif(S0 = S0[1], r = r, 
               sigma = sigma[1], Time = Time, min = x_min, max=x_max)
      S2[i] <- generateTermGeomBMPathsUnif(S0 = S0[2], r = r,
               sigma = sigma[2], Time = Time, min = y_min, max=y_max)
      call_prices[i] <- exp(-r*Time)*max( 0.5*(S1[i] + S2[i]) - K, 0)
    }
    bin_price[k,j] <- mean(call_prices)
    bin_var[k,j] <- var(call_prices)  
  }
}

call_price <- mean(bin_price)
se_call <- sqrt(sum(bin_var)/(num_bins*sampl_size))
comp_df$Call_Price[2] <- call_price
comp_df$SE_Call_Price[2] <- se_call
cat("Call Price and Standard Error\n")
cat("Call Price: ", call_price, "\n")
cat("Standard Error: ", se_call)
```

## 1(c)
```{r 1(c), dependson="paramsq1", cache=TRUE}
num_bins <- 250
bin_size <- 4
nu <- c(1/sqrt(2), 1/sqrt(2))


bin_price <- numeric()
bin_var <- numeric()
for(i in 1:num_bins){
  call_prices <- numeric()
  for(j in 1:bin_size){
    x_min <- (i-1)/num_bins
    x_max <- i/num_bins
    # generate the uniform random number
    x <- qnorm(runif(1, min = x_min, max = x_max))
    mean <- nu*x
    sigmaMat <- diag(2) - nu%*%t(nu)
    z <- mvrnorm(n = 1, mu = mean, Sigma = sigmaMat)
    S1 <- S0[1]*exp((r-0.5*sigma[1]^2)*Time+sigma[1]*z[1]*sqrt(Time))
    S2 <- S0[2]*exp((r-0.5*sigma[2]^2)*Time+sigma[2]*z[2]*sqrt(Time))
    call_prices[j] <- exp(-r*Time)*max(0.5*(S1 + S2) - K, 0)  
  }
  bin_price[i] <- mean(call_prices)
  bin_var[i] <- var(call_prices)
}

call_price <- mean(bin_price)
se_call <- sqrt(sum(bin_var)/(num_bins*sampl_size))
comp_df$Call_Price[3] <- call_price
comp_df$SE_Call_Price[3] <- se_call
cat("Call Prices and Standard Errors\n")
cat("Call Price: ", call_price, "\n")
cat("Standard Error: ", se_call)
knitr::kable(comp_df)
```

We observe that the standard error reduces significantly(becomes ~1/4 th) by using the bivariate stratification. The conditioning on projection method gives significant reduction in the standard error as compared to both the methods.

## Question 2
```{r paramsq2}
S0 <- 50
mu <- 0.15
r <- 0.1
sig <- 0.25
K <- 50
sampl_size <- 1000
N <- 30
Time <- 0.25
```
## 2(a)
## standard Monte Carlo
```{r 2(a), cache=TRUE}
opt_prices <- numeric()
for(i in 1:sampl_size){
  stockPrice <- generateGeomBMPaths(S0 = S0, r = r, sigma = sig,
                Time = Time, N = 30)
  opt_prices[i] <- exp(-r*Time)*max(max(stockPrice)-K,0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)

```
## Brownian Bridge
```{r, cache=TRUE}
deltaTime <- Time/N
opt_prices <- numeric()
for(i in 1:sampl_size){
 stockPrice <- numeric()
 maxPrice <- S0
 stockPrice[1] <- S0 + r*S0*deltaTime + sig*S0*sqrt(deltaTime)*rnorm(1) 
 b <- (stockPrice[1]-S0)/(sig*S0)
 maxPrice <- S0 + sig*S0*0.5*(b + sqrt(b^2 - 2*deltaTime*log(1-runif(1))))
 
 for(k in 2:N){
   stockPrice[k] <- stockPrice[k-1] + r*stockPrice[k-1]*deltaTime + 
                    sig*stockPrice[k-1]*sqrt(deltaTime)*rnorm(1)
   b <- (stockPrice[k] - stockPrice[k-1])/(sig*stockPrice[k-1])
   maxPrice <- max(stockPrice[k-1] + sig*stockPrice[k-1]*
              0.5*(b + sqrt(b^2 - 2*deltaTime*log(1-runif(1)))), maxPrice)
 }
 opt_prices[i] <- exp(-r*Time)*max(maxPrice - K, 0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```

The option price obtained using Brownian bridge is higher than the one obtained using Standard Monte Carlo technique.

## 2(b)
## using the final price as the strike price
```{r, cache=TRUE}
opt_prices <- numeric()
for(i in 1:sampl_size){
  stockPrice <- generateGeomBMPaths(S0 = S0, r = r, sigma = sig,
                Time = Time, N = 30)
  opt_prices[i] <- exp(-r*Time)*max(max(stockPrice)-stockPrice[N],0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```
## Brownian Bridge
```{r, cache=TRUE}
deltaTime <- Time/N
opt_prices <- numeric()
for(i in 1:sampl_size){
 stockPrice <- numeric()
 maxPrice <- S0
 stockPrice[1] <- S0 + r*S0*deltaTime + sig*S0*sqrt(deltaTime)*rnorm(1) 
 b <- (stockPrice[1]-S0)/(sig*S0)
 maxPrice <- S0 + sig*S0*0.5*(b + sqrt(b^2 - 2*deltaTime*log(1-runif(1))))
 
 for(k in 2:N){
   stockPrice[k] <- stockPrice[k-1] + r*stockPrice[k-1]*deltaTime + 
                    sig*stockPrice[k-1]*sqrt(deltaTime)*rnorm(1)
   b <- (stockPrice[k] - stockPrice[k-1])/(sig*stockPrice[k-1])
   maxPrice <- max(stockPrice[k-1] + sig*stockPrice[k-1]*
              0.5*(b + sqrt(b^2 - 2*deltaTime*log(1-runif(1)))), maxPrice)
 }
 opt_prices[i] <- exp(-r*Time)*max(maxPrice - stockPrice[N], 0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```

Compared to the values in the table 1, the options prices are lower as expected as the expected terminal stock price is $S_0e^{rT}$. Also, The option price obtained using Brownian bridge is higher than the one obtained using Standard Monte Carlo technique.


## 2(c)
## knock-out options
```{r params2c, cache=TRUE}
S0 <- 50
r <- 0.1
sig <- 0.5
K <- 50
H <- 45
N <- 30
sampl_size <- 100000
```
## standard monte carlo
```{r, dependson="params2c",cache=FALSE}
opt_prices <- numeric()
for(i in 1:sampl_size){
  stockPrice <- generateGeomBMPaths(S0 = S0, r = r, sigma = sig,
                                    Time = Time, N = 30)
  barr_crossed <- ifelse(min(stockPrice) < H, 1, 0)
  opt_prices[i] <- exp(-r*Time)*(1-barr_crossed)*max(stockPrice[N]-K,0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```
## Brownian Bridge
```{r, cache=TRUE}
deltaTime <- Time/N
opt_prices <- numeric()
for(i in 1:sampl_size){
 stockPrice <- numeric()
 minPrice <- 0
 stockPrice[1] <- S0 + r*S0*deltaTime + sig*S0*sqrt(deltaTime)*rnorm(1) 
 b <- (stockPrice[1]-S0)/(sig*S0)
 minPrice <- S0 + sig*S0*0.5*(b - sqrt(b^2 - 2*deltaTime*log(1-runif(1))))
 
 for(k in 2:N){
   stockPrice[k] <- stockPrice[k-1] + r*stockPrice[k-1]*deltaTime + 
                    sig*stockPrice[k-1]*sqrt(deltaTime)*rnorm(1)
   b <- (stockPrice[k] - stockPrice[k-1])/(sig*stockPrice[k-1])
   minPrice <- min(stockPrice[k-1] + sig*stockPrice[k-1]*
              0.5*(b - sqrt(b^2 - 2*deltaTime*log(1-runif(1)))), minPrice)
 }
 barr_crossed <- ifelse(minPrice < H, 1, 0)
 opt_prices[i] <- exp(-r*Time)*(1-barr_crossed)*max(stockPrice[N] - K, 0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```

The option price obtained using Brownian bridge is lower than the one obtained using Standard Monte Carlo technique.

## Question 3
```{r paramsq3}
S0 <- c(100, 100)
Time <- 0.2
r <- 0.1
sigma <- c(0.3, 0.3)
rho <- 0.5
K <- 100
H <- 95
N <- 50
sampl_size <- 100000
deltaTime <- Time/N
```

##3(a)
## Standard Monte Carlo 
```{r, dependson="paramsq3", cache=TRUE}
opt_prices <- numeric()
for(i in 1:sampl_size){
  S1 <- numeric()
  S2 <- numeric()
  mu <- c(0,0)
  sigmaMat <- matrix(data=c(1,rho,rho,1), nrow = 2, ncol = 2)
  z <- mvrnorm(n = 1, mu = mu, Sigma = sigmaMat)
  S1[1] <- S0[1] + r*S0[1]*deltaTime + sigma[1]*S0[1]*sqrt(deltaTime)*z[1]
  S2[1] <- S0[2] + r*S0[2]*deltaTime + sigma[2]*S0[2]*sqrt(deltaTime)*z[2]
  
  for(k in 2:N){
    z <- mvrnorm(n = 1, mu = mu, Sigma = sigmaMat)
    S1[k] <- S1[k-1] + r*S1[k-1]*deltaTime + sigma[1]*S1[k-1]*sqrt(deltaTime)*z[1]
    S2[k] <- S2[k-1] + r*S2[k-1]*deltaTime + sigma[2]*S2[k-1]*sqrt(deltaTime)*z[2]
  }
  barr_crossed <- ifelse(min(S2)<H, 1, 0)
  opt_prices[i] <- exp(-r*Time)*(1-barr_crossed)*max(S1[N]-K,0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```

##3(b)
## Brownian Bridge
```{r, dependson="paramsq3", cache=TRUE}
opt_prices <- numeric()
for(i in 1:sampl_size){
  S1 <- numeric()
  S2 <- numeric()
  mu <- c(0,0)
  sigmaMat <- matrix(data=c(1,rho,rho,1), nrow = 2, ncol = 2)
  z <- mvrnorm(n = 1, mu = mu, Sigma = sigmaMat)
  S1[1] <- S0[1] + r*S0[1]*deltaTime + sigma[1]*S0[1]*sqrt(deltaTime)*z[1]
  S2[1] <- S0[2] + r*S0[2]*deltaTime + sigma[2]*S0[2]*sqrt(deltaTime)*z[2]
  minPrice <- 0
  b <- (S2[1]-S0[2])/(sigma[2]*S0[2])
  minPrice <- S0[2] + sigma[2]*S0[2]*0.5*(b - sqrt(b^2 - 2*deltaTime*
                                                     log(1-runif(1))))
  for(k in 2:N){
    z <- mvrnorm(n = 1, mu = mu, Sigma = sigmaMat)
    S1[k] <- S1[k-1] + r*S1[k-1]*deltaTime + sigma[1]*S1[k-1]*sqrt(deltaTime)*z[1]
    S2[k] <- S2[k-1] + r*S2[k-1]*deltaTime + sigma[2]*S2[k-1]*sqrt(deltaTime)*z[2]
    b <- (S2[k] - S2[k-1])/(sigma[2]*S2[k-1])
    minPrice <- min(S2[k-1] + sigma[2]*S2[k-1]*0.5*(b - sqrt(b^2 - 2*deltaTime*
                log(1-runif(1)))), minPrice)
  }
  barr_crossed <- ifelse(minPrice < H, 1, 0)
  opt_prices[i] <- exp(-r*Time)*(1-barr_crossed)*max(S1[N]-K, 0)
}

opt_price <- mean(opt_prices)
se_opt_price <- sqrt(sum((opt_prices - opt_price)^2)/(sampl_size*(sampl_size-1)))

cat("Options Price and Standard Error\n")
cat("Option Price: ", opt_price, "\n")
cat("Standard Error: ", se_opt_price)
```


# Question 4
```{r paramsq4}
N <- 5
Time <- 5
r <- 0.04
RRate <- 0.35
s <- 0.01
rhos <- seq(0,1,by = 0.2)
sampl_size <- 100000

# lambda
lam <- s/(1-RRate)

```

```{r, dependson="paramsq4", cache=TRUE}
res_df <- data.frame(matrix(nrow = length(rhos), ncol = 11))
colnames(res_df) <- c("Corr","Ftd", "FtdSE","2td", "2tdSE","3td", "3tdSE",
                      "4td", "4tdSE","5td", "5tdSE")
res_df$Corr <- rhos

for(j in 1:nrow(res_df)){
  rho <- res_df$Corr[j]
  sigmaMat <- diag(N)
  sigmaMat[lower.tri(sigmaMat)] <- rep(rho,sum(lower.tri(sigmaMat)))
  sigmaMat[upper.tri(sigmaMat)] <- t(sigmaMat[lower.tri(sigmaMat)])
  # generate Gaussian coupla
  if(rho!=1){
    A <- chol(sigmaMat)
  }
  prices <- matrix(nrow = sampl_size, ncol = N)
  for(i in 1:sampl_size){
    def_time <- numeric()
    if(rho == 1){
      def_time <- -log(rep(runif(1),N))/lam
    }else{
      Z <- rnorm(N)
      Y <- t(A)%*%Z
      U <- pnorm(Y)
      def_time <- -log(U)/lam
    }
    # count the number of bonds that defaulted
    num_def <- sum(def_time < Time)
    prices[i,] <- exp(-r*Time)*(1-RRate)*ifelse(num_def >= seq(1,5), 1, 0)
  }
  res_df[j, seq(2,10,2)] <- round(apply(prices, MARGIN = 2, mean),6)
  res_df[j, seq(3,11,2)] <- round(apply(prices, MARGIN = 2, function(x)
                            {calculateMeanSE(x)}),6)
}

knitr::kable(res_df)
```

