---
title: "Time Series HW5"
author: "Adil Hayat"
date: "23 April 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vars)
library(dlm)
```

## Question 1
```{r Problem 6}
data(Tbrate,package="Ecdat")
# r = the 91-day Treasury bill rate
# y = the log of real GDP
# pi = the inflation rate
del_dat = diff(Tbrate)
var1 = ar(del_dat,order.max=4,aic=T)
var1
acf(var1$resid[-1,])
```

Problem 6 For this problem, use the notation of equation (10.7) with q = 0.

_(a) What is p and what are the estimates $\Phi_1$; : : : ;$\Phi_p$?_

The value of p being used is `r var1$order`. The value of $\Phi_1$ is given in the table below
`r ph <- as.data.frame(matrix(var1$ar,nrow = 3, ncol = 3))
colnames(ph) <- c('r','y','pi')
rownames(ph) <- c('r','y','pi')
knitr::kable(ph)`

_(b) What is the estimated covariance matrix of $\epsilon_t$?_
The value of the estimated covariance matrix is given below:
`r ec <- as.data.frame(matrix(var1$var.pred,nrow = 3, ncol = 3))
colnames(ec) <- c('r','y','pi')
rownames(ec) <- c('r','y','pi')
knitr::kable(ec)`

_(c) If the model fits adequately, then there should be no residual auto- or cross-correlation. Do you believe that the model does fit adequately?_

We observe that the there is no residual auto or cross-correlation. So, we can conculde that the model fits adequately.

## Question 2
```{r, warning=FALSE, cache=TRUE}
dat <- read.table("m-ppiaco4709.txt")
colnames(dat) <- c("Year","Month","Day","PPI")
z <- diff(log(dat$PPI))

buildAR = function(x)
{
   FF = matrix(c(1,0,0), nrow=1)
  
   GG = matrix(c(x[1],1,0,x[2],0,1,x[3],0,0),nrow = 3, ncol=3)

   V = diag(1) * exp(x[4])

   W = matrix(c(1,rep(0,8)),nrow = 3, ncol=3) * exp(x[5])
   
   m0 = c(0,0,0)

   C0 = matrix(solve(diag(9) - GG %x% t(GG) ) %*% matrix(W,ncol=1), ncol=3)

   return(list(m0 = m0, C0 = C0, FF = FF, GG = GG, 
       V = V, W = W))
}

holdfit = dlmMLE(z,c(0,0,0,1,1),buildAR)

# Model Parameters
param_df <- data.frame(matrix(nrow=5, ncol=2))
colnames(param_df) <- c("Parameter", "Values")
param_df$Parameter <- c("Phi1","Phi2","Phi3","var_epsilon", "var_a")
param_df$Values <- c(holdfit$par[1], holdfit$par[2], holdfit$par[3], 
                     exp(holdfit$par[4]), exp(holdfit$par[5]))

filtout = dlmFilter(z,buildAR(holdfit$par))
smoothout = dlmSmooth(z,buildAR(holdfit$par))

```

**(a) Write this model in state space form.**

$$\mathbf{X_t} = \mathbf{G}\mathbf{X_{t-1}} + \mathbf{w_t}$$
$$Y_t = \mathbf{F}\mathbf{X_{t}} + v_t$$
$$\mathbf{G}=\begin{bmatrix}
\phi_{1} & \phi_{2} & \phi_{3}\\
1 & 0 & 0\\
0 & 1& 0\\
\end{bmatrix}$$

$$\mathbf{F} = [1, 0, 0]$$
$$\mathbf{X_t} = [X_t, X_{t-1}, X_{t-2}]^T$$
$$v_t = \epsilon_t$$
$$\mathbf{w_t} = [a_t, 0, 0]^T$$
$$\mathbf{V} = [\sigma_\epsilon^2]$$
$$\mathbf{W}=\begin{bmatrix}
\sigma_a^2 & 0 & 0\\
0 & 0 & 0\\
0 & 0& 0\\
\end{bmatrix}$$

 


**(b) The file m-ppiaco4709.txt (available on Blackboard) contains year, month, day, and U.S. producer price index (PPI) from January 1947 to November 2009. The index is for all commodities, and not seasonally adjusted. Let $z_t=log(Z_t)-log(Z_{t-1})$ where $Z_t$ is the observed monthly PPI. It turns out that an AR(3) model is adequate for $z_t$ if the minor seasonal dependence is ignored. Let yt be the sample mean-corrected series of $z_t$. Suppose that yt has independent measurement errors so that $y_t = x_t + \epsilon_t$ where xt is a mean-zero AR(3) process and V ($\epsilon_t$) = $\sigma_t^2$. Use the state space model to estimate the parameters. Write down the fitted model. Also, create a plot showing the filtered estimates of xt as a function of t.**

```{r}
knitr::kable(param_df)
plot(ts(z),lwd=1,cex.axis=1.3,cex.lab=1.3,ylab="Log Diff PPI",
     main="Actual vs Filtered Values")
points(ts(filtout$m),lwd=1,col=4,lty=1,type="p", pch="+")
legend("topright",c("Actual", "Filtered"),col=c("black", "blue"),
       lty=c(1,NA), pch=c(NA,"+"), bg='white')
plot(ts(filtout$m[,1]),lwd=1,col=4,lty=1,type="l", main="Filtered Estimates")
```

