---
title: "Time Series HW3"
author: "Adil Hayat"
date: "12 April 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(quantmod))
suppressPackageStartupMessages(library(tseries))
suppressPackageStartupMessages(library(forecast))
```


```{r, cache=TRUE}
TB3MS <- getSymbols("TB3MS", src="FRED", auto.assign = FALSE)
TB3MS_sub <- TB3MS["1990-01-01/2012-12-31"]
plot(TB3MS_sub, ylab="3-month T-Bill Rate", main="3-month Treasury Bill Rate")
```

The series does not seem stationary as the time series doesn't seem to fluctuate around a particular level.

## Question 2(b)
## Taking the differences
```{r}
TB3MS_diff1 <- diff(TB3MS_sub,lag = 1, differences = 1)[-1]
TB3MS_diff2 <- diff(TB3MS_sub,lag = 1, differences = 2)[-1:-2]
plot(TB3MS_diff1, main="1st Order Differenced T-Bill")
par(mfrow=c(1,2))
acf(TB3MS_diff1, main = "1st Order Differencing")
acf(TB3MS_diff2, main = "2nd Order Differencing")
adf.test(TB3MS_diff1)
```

The difference series looks staionary as it fluctuates around 0 and there is no clear pattern in the variance across time. The acf plot with a difference of 1 looks reasonable while if we take the difference twice we get an acf plot with significant negative correlation at lag 1 suggesting that we have applied differencing more than that is needed.
ADF test also suggests that the 1st order differenced series is stationary.

## Question 2(c)
```{r}
holdfit <- auto.arima(TB3MS_sub, d=1, approximation = FALSE, step = FALSE, ic="aicc")
holdfit
```

## Question 2(d)
```{r}
acf(holdfit$residuals,main= "ACF of the residuals", na.action = na.omit)
Box.test(holdfit$residuals, fitdf = 3, lag=20, type="Ljung-Box")
```

The model seems to be a reasonable fit for the data as we observe no significant auto-correlation at lower lags and there is no pattern to the lags as well. The observed auto-correlation at higher lags can just be an artifact of way hypothesis testing is designed. From the Box Test, we observe that there is some residual correlation at some lags.

## Question 2(e)
```{r}
holdpreds <- predict(holdfit, n.ahead = 3)
lower_int <- holdpreds$pred - 1.96*holdpreds$se
upper_int <- holdpreds$pred + 1.96*holdpreds$se
cat("Upper Interval:", as.double(upper_int),"\n")
cat("Lower Interval:", as.double(lower_int))
```

# Question 2(f)
Yes, the intervals include the observed values as the intervals shown for the data points as shown above.
The actual values for 3-month T-Bill rate were **0.07, 0.1, 0.09**.
